---
title: "PGH Accessibility"
author: "Adam Peterson"
date: "11/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sf)
library(mapview)
mapviewOptions(platform ='leafgl')
library(xml2)
library(mapsapi)
set.seed(34231)
## loads in google API key
source("keyfile.R")
url <- "https://data.wprdc.org/dataset/ece64ad3-05eb-46dd-ba38-c83b5373812f/resource/3f40b94b-4ac4-48f1-8c61-8439d2d2f420/download/wprdc_stop_data.csv"
```




* Sample points near bus routes in each neighborhood.
* calculate trip distance to calculate accessibility index
by looking at trip time/distance to: 

1. nearest grocery store
2. downtown
3. commute (using sample)
4. other areas of interest (e.g. Univ., hospital, etc.)

By mode of transit 

* bike (owned vs. share)
* bus
* car
* walk
* scooter 


Taking into account

* Level of stress (bike)
* park/traffic hassle
* scooter access (if data)
* COST!! 

```{r datasets}
bdf <- read_csv(url) %>% 
    st_as_sf(.,coords=c("longitude","latitude")) %>% 
    st_set_crs(4326)


brdf <- read_sf("~/Documents/CityData/Burgh/paac_routes_1909/PAAC_Routes_1909.shp")

lots <- read_sf("~/Documents/CityData/Burgh/Zoning/Zoning.shp")

hoods <- read_sf("~/Documents/CityData/Burgh/Neighborhoods/Neighborhoods_.shp") %>%
    select(objectid,hood) %>%
    st_transform(4326)

```


```{r}
## Downloaded from https://www.pasda.psu.edu/uci/DataSummary.aspx?dataset=1214
pcls <- read_sf("~/Documents/CityData/Burgh/AlleghenyCounty_Parcels202111/AlleghenyCounty_Parcels202111.shp")
```


```{r}
resid <- lots %>% filter(str_detect(legendtype,"Neighborhood|Residential|Public"))
residpcls <- pcls %>% 
    mutate(valid = map_lgl(geometry,st_is_valid)) %>% 
    filter(valid == TRUE) %>% 
    st_filter(resid,join = st_contains) %>% 
    st_transform(st_crs(hoods)) %>% 
    st_join(hoods) %>% 
    filter(hood=="Shadyside") %>% 
    st_centroid()
```


```{r}
shdysample <- residpcls %>% st_sample(size = 100) %>% st_set_crs(4326) 
origin <- as.matrix(shdysample[[1]])
destinations <- c("Shadyside UPMC",
                  "Giant Eagle",
                  "Point State Park",
                  "Schenley Park")

query_google <- function(origins,
                         destinations,
                         key,
                         traffic_model = "best_guess",
                         depart_time = as.POSIXct("2021-11-29 08:00:00 EST")){
    
    query_modes_and_process <- function(origins,
                                  destinations,
                                  mode,
                                  traffic_model,
                                  depart_time){
        
        process <- function(query_result,data_val = "distance_m"){
            out <- as_tibble(mp_get_matrix(as_xml_document(query_result), 
                             value = data_val),
                             rownames = "Origin") %>% 
             mutate(trip_ix = 1:n()) %>% 
             gather(everything(),-trip_ix,-Origin,
                    key = "Destination",value = !!data_val) %>% 
             mutate(trip_ix2 = 1:n()) %>% 
                gather({{data_val}},key = "TripDataType", value = "TripData")
            return(out)
        }
        
        foo <- mp_matrix(origins = origins,
                       destinations = destinations,
                       mode = mode,
                       key = key,
                       departure_time = depart_time,
                       traffic_model = traffic_model,
                       quiet = TRUE)
        out <- map_dfr(c("distance_m","duration_s","duration_in_traffic_s"), 
                       function(data_val) process(foo,data_val) %>%
                           mutate(transit_mode = mode))
        return(out)
        
    }
    
    out <- map_dfr(c("driving","transit","walking","bicycling"), function(mode){
                query_modes_and_process(origins,destinations,mode,
                                        traffic_model,depart_time)})
    origin_latlong <- tibble(trip_ix = 1:nrow(origins),
                             lat = origins[,2],
                             long = origins[,1])
    out <- out %>% left_join(origin_latlong,by="trip_ix")
    return(out)
}
```


```{r,eval = FALSE}
shxdf <- query_google(origins = origin,
                      destinations = destinations,
                      key = key)
```


```{r}
shxdf <- read_csv("~/Documents/CityData/Burgh/Google_Accessibility/shadyside.csv")
shxdf %>% mutate(TripData = case_when(TripDataType == "distance_m" ~ TripData / 1000,
                                      TripDataType == "duration_s" ~ TripData / 60,
                                      TRUE ~ TripData / 60)) %>% 
    group_by(Destination,transit_mode,TripDataType) %>% 
    summarise(avg = mean(TripData),
              sprd = var(TripData))
```


## r5r 

```{r}
options(java.parameters = "-Xmx2G")
library(r5r)
origins_r5r <- shdysample %>% st_cast(.,"POINT") %>% 
    st_sf() %>% mutate(id=1:n()) 
destinations_r5r <- c(st_point(c(-79.919479,40.459080)),
                      st_point(c(-79.921059,40.458851)),
                      st_point(c(-79.928703,40.458569)) 
) %>% 
    st_sfc() %>% 
    st_cast(.,"POINT") %>% 
    st_sf() %>% 
    mutate(id=c("Trader Joes","Giant Eagle","Whole Foods"))
```


```{r}
r5r_core <- setup_r5(data_path = "~/Documents/CityData/Burgh/TransitNetworkData/")
```